\documentclass[lettersize,journal]{IEEEtran}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{booktabs}
\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore}

\begin{document}

\title{Precise Visual Servoing for Suction Cup Grippers Using Foundation Model Segmentation}

\author{Akanksha Singal,~\IEEEmembership{Robotics Institute,~CMU}
\thanks{This work was carried out as part of an independent study in the Master of Science
in Computer Vision program at Carnegie Mellon University's Robotics Institute under the
supervision of Prof.\ Jeffrey Ichnowski, Momentum Robotics Lab.}}

\markboth{IEEE Robotics and Automation Letters}%
{Singal: Precise Visual Servoing for Suction Cup Grippers}

\maketitle

% =========================================================================
\begin{abstract}
TODO: Write abstract after experiments are complete.
\end{abstract}

\begin{IEEEkeywords}
Visual servoing, suction grasping, image-based visual servoing, foundation models,
SAM3, object segmentation, xArm, ZED Mini.
\end{IEEEkeywords}


% =========================================================================
\section{Introduction}
\label{sec:intro}

TODO: Write introduction after results are available.


% =========================================================================
\section{Related Work}
\label{sec:related}

TODO: Write related work section.


% =========================================================================
\section{System Overview}
\label{sec:system}

The proposed system performs image-based visual servoing (IBVS) using a
6-DOF xArm manipulator equipped with a ZED Mini stereo camera mounted
rigidly on the tool flange in an eye-in-hand configuration. The camera
captures frames at HD720 resolution ($1280 \times 720$ pixels). The system
operates in two sequential phases: (i) a far-range approach phase driven by
SAM3 open-vocabulary segmentation, and (ii) a close-range last-mile phase that
switches to a multi-signal fusion estimator when the target occupies a
sufficiently large portion of the image and is within the reliable stereo
depth range.

Robot motion commands are issued via the xArm Python SDK at a maximum
rate of one command every $0.3$\,s, with a TCP speed limit of
$150$\,mm\,s$^{-1}$ and an acceleration limit of $500$\,mm\,s$^{-2}$.


% =========================================================================
\section{Methodology}
\label{sec:method}

% -------------------------------------------------------------------------
\subsection{Image-Based Visual Servo Control Law}
\label{sec:ibvs}

The system employs a decoupled image-based proportional controller.
Let $(u_c, v_c)$ denote the pixel coordinates of the estimated target
centroid and $(W/2,\,H/2)$ the image center, where $W=1280$ and $H=720$.
The image-plane error is
\begin{equation}
  e_x = u_c - W/2, \qquad e_y = v_c - H/2.
  \label{eq:pixerr}
\end{equation}
The robot end-effector (EE) displacement commanded at each servo step is:
\begin{align}
  \Delta x &= 3.0\ \text{mm} \quad \text{(constant forward approach)},
  \label{eq:dx}\\
  \begin{bmatrix}\Delta y \\ \Delta z\end{bmatrix}
  &= -\alpha\,\hat{\mathbf{J}}^{+}_{yz}
     \begin{bmatrix}e_x \\ e_y\end{bmatrix},
  \label{eq:dyz}
\end{align}
where $\alpha = 0.5$ is a fixed control gain and $\hat{\mathbf{J}}^{+}_{yz}$
is the Moore--Penrose pseudoinverse of the estimated $2\times2$ image Jacobian
(Section~\ref{sec:jacobian}). The lateral commands are saturated to
$|\Delta y|, |\Delta z| \leq 12.0$\,mm per step.
The lateral correction in~\eqref{eq:dyz} is applied only when the radial
pixel error $\lVert\mathbf{e}\rVert_2 > 15$\,px (the dead zone); within this
dead zone, $\Delta y = \Delta z = 0$ while the forward approach~\eqref{eq:dx}
continues.  A safety floor clamps the Z coordinate of the EE to
$z_{\text{EE}} \geq 0$\,mm in the robot base frame.

% -------------------------------------------------------------------------
\subsection{Online Jacobian Calibration}
\label{sec:jacobian}

The image Jacobian $\hat{\mathbf{J}}_{yz}$ relates a millimetre-scale robot
translation in the lateral plane to the resulting pixel displacement in the
image, and is calibrated once at the start of each trial via small
perturbations. For each lateral axis $k \in \{Y, Z\}$:
\begin{enumerate}
  \item Capture frame $\mathbf{I}_{\text{before}}$ at the current EE pose.
  \item Translate the EE by $\delta = 8.0$\,mm along axis $k$ at
        $30$\,mm\,s$^{-1}$ and $100$\,mm\,s$^{-2}$.
  \item Wait $1.5$\,s for vibration to settle.
  \item Capture frame $\mathbf{I}_{\text{after}}$.
  \item Return the EE to the original pose.
  \item Compute the median optical flow displacement $(dp_x, dp_y)$ between
        $\mathbf{I}_{\text{before}}$ and $\mathbf{I}_{\text{after}}$ using
        pyramidal Lucas--Kanade tracking on up to 300 Shi--Tomasi corner
        features (minimum distance 8\,px, quality level 0.01).
  \item Set the $k$-th column of $\hat{\mathbf{J}}_{yz}$ to
        $[dp_x/\delta,\ dp_y/\delta]^T$.
\end{enumerate}
If the resulting Jacobian has rank zero (optical flow failure), the system
falls back to approach-only mode ($\Delta y = \Delta z = 0$). If rank is 1,
partial lateral control is available via the pseudoinverse.

% -------------------------------------------------------------------------
\subsection{Far-Range Phase: SAM3 Segmentation and Target Tracking}
\label{sec:sam3}

During the far-range approach, the target is detected and segmented on every
camera frame using SAM3 (Segment Anything Model~3), loaded from the
\texttt{facebook/sam3} checkpoint via the HuggingFace Transformers library.
SAM3 performs Promptable Concept Segmentation (PCS), jointly detecting and
segmenting all instances of a concept in a single forward pass from the text
prompt ``\texttt{box}''.

An instance is accepted if its score exceeds $\tau_s = 0.50$ and its mask
probability exceeds $\tau_m = 0.50$. If no detection passes these thresholds,
a permissive retry is executed with $\tau_s = 0.30$ and $\tau_m = 0.40$.
Non-maximum suppression (IoU threshold 0.45) is applied to the resulting set
of candidate bounding boxes, after which instances with bounding-box area
below $0.3\%$ or above $75\%$ of the image area, or with aspect ratio
exceeding 8:1, are discarded.

\subsubsection{Reference-Guided Disambiguation}
\label{sec:disambiguation}

When SAM3 returns multiple instances, the target box is selected by a
composite score that combines visual similarity to a pre-captured reference
image, positional preference for boxes appearing near the top of the image
(which corresponds to the top of a stack), SAM3 detection confidence, and
temporal proximity to the prior centroid:
\begin{equation}
  s = 0.50\,s_{\text{sim}} + 0.15\,s_{\text{pos}} + 0.10\,s_{\text{conf}}
      + 0.25\,s_{\text{prox}},
  \label{eq:disambig}
\end{equation}
where:
\begin{itemize}
  \item $s_{\text{sim}}$ is the cosine similarity between the L2-normalised
        512-dimensional ResNet-18 feature vector (penultimate
        average-pooling layer, ImageNet pretrained) extracted from the
        candidate crop and a reference feature vector computed offline from
        a single reference image of the target box type;
  \item $s_{\text{pos}} = 1 - v_c / H$ favours instances whose vertical
        centre is higher in the image (i.e., closer to the top face of a
        stacked box);
  \item $s_{\text{conf}}$ is the raw SAM3 instance confidence score;
  \item $s_{\text{prox}} = \max(0,\, 1 - d\,/\,(0.25\max(H,W)))$ is the
        normalised proximity of the candidate centre to the centroid
        detected in the previous frame, providing temporal continuity.
\end{itemize}
When no reference image is provided, the weights collapse to
$s_{\text{pos}}:s_{\text{conf}}:s_{\text{prox}} = 0.50:0.25:0.25$.

\subsubsection{Grasp-Point Estimation}
\label{sec:centroid}

The target centroid --- used as the error signal $\mathbf{e}$ in
\eqref{eq:pixerr} --- is the rotated-rectangle centre of the largest
contour of the morphologically closed binary mask (structuring element:
$15\times15$ rectangle, 2 iterations). If the rectangle centre falls outside
the mask, the convex hull centroid is used as a fallback; if that also fails,
the standard image-moment centroid is used.

\subsubsection{Mask Stability and IoU-Drift Detection}
\label{sec:tracker}

Frame-to-frame mask stability is tracked via the intersection-over-union (IoU)
between consecutive masks. An IoU below 0.35 on a frame that used the
prior bounding box as a SAM3 prompt forces a full open-vocabulary redetection
on the next frame. An anchor mask is locked when IoU $\geq 0.82$ is sustained
for five consecutive frames; the anchor is used as a stable spatial reference
for redetection after tracking loss.

A centroid jump exceeding 80\,px between consecutive servo frames is treated
as a spurious detection and the step is skipped without issuing a robot command.

% -------------------------------------------------------------------------
\subsection{Last-Mile Phase: Multi-Signal Fusion}
\label{sec:lastmile}

As the camera approaches the target, the bounding box eventually exceeds the
image border, causing the mask centroid to drift toward whichever side of the
object remains visible. To maintain accurate target tracking under this
partial-framing condition, the system switches to a multi-signal fusion
estimator implemented in \texttt{servo\_lastmile.py}. The estimator operates
as a discrete state machine with four states: \textsc{far}, \textsc{lock},
\textsc{near}, and \textsc{terminal}.

\subsubsection{FAR State}

In the \textsc{far} state, the system runs the same SAM3 pipeline described
in Section~\ref{sec:sam3} on every frame and evaluates the lock-transition
criterion on each detection.

\subsubsection{Lock Transition}

A transition from \textsc{far} to \textsc{lock} is triggered when all of the
following conditions hold for three consecutive frames:
\begin{itemize}
  \item The bounding box does not touch the image border (margin: 8\,px, and
        5\% of the corresponding image dimension);
  \item The bounding box occupies between $5\%$ and $25\%$ of the total image
        area;
  \item The depth at the mask centroid, $z_{\text{lock}}$, satisfies
        $150 \leq z_{\text{lock}} \leq 400$\,mm.
\end{itemize}
Upon lock, a \texttt{LockState} object is captured containing: the binary
mask, the 2-D centroid, the bounding box, an image crop (bounding box padded
by $10\%$ on each side), the depth $z_{\text{lock}}$, the 4$\times$4
homogeneous EE pose $\mathbf{T}_{\text{base,ee}}$, and the 3-D centroid
in the robot base frame $\mathbf{p}_{\text{base}}$.

The locked 3-D centroid is computed from the camera-frame coordinates:
\begin{equation}
  \mathbf{p}_{\text{cam}} =
    \begin{bmatrix}
      (u_c - c_x)\,z_{\text{lock}} / f_x \\
      (v_c - c_y)\,z_{\text{lock}} / f_y \\
      z_{\text{lock}}
    \end{bmatrix},
  \label{eq:backproject}
\end{equation}
\begin{equation}
  \mathbf{p}_{\text{base}} =
    \mathbf{T}_{\text{base,ee}}\,\mathbf{T}_{\text{ee,cam}}\,
    \tilde{\mathbf{p}}_{\text{cam}},
  \label{eq:transform}
\end{equation}
where $(f_x, f_y, c_x, c_y)$ are the ZED Mini camera intrinsics,
$\mathbf{T}_{\text{ee,cam}}$ is the hand-eye calibration matrix (camera pose
in the EE frame), and $\tilde{\mathbf{p}}_{\text{cam}}$ denotes the
homogeneous coordinate of $\mathbf{p}_{\text{cam}}$.

CoTracker3 is also initialised at lock with 80 uniformly sampled points
inside the lock mask, using a fixed random seed.

\subsubsection{NEAR State: Four-Signal Fusion}

In the \textsc{near} state, four parallel signals produce candidate centroid
estimates that are fused via a weighted geometric median:

\textbf{Signal A --- Locked-3D Projection.}
The frozen base-frame centroid $\mathbf{p}_{\text{base}}$ is re-projected
into the current image frame using the live EE pose:
\begin{equation}
  \tilde{\mathbf{p}}_{\text{cam}}^{\text{now}}
  = \bigl(\mathbf{T}_{\text{base,ee}}^{\text{now}}\,
           \mathbf{T}_{\text{ee,cam}}\bigr)^{-1}
    \tilde{\mathbf{p}}_{\text{base}},
\end{equation}
followed by pinhole projection. Signal~A is assigned weight $w_A = 1.0$
when a hand-eye calibration is available, and $w_A = 0.0$ otherwise.

\textbf{Signal B --- CoTracker3 with LK Fallback.}
CoTracker3~Online (loaded from \texttt{facebookresearch/co-tracker} via
\texttt{torch.hub}) propagates the 80 initialisation points forward in time.
Points with predicted visibility below 0.5 are excluded; the signal centroid
is the unweighted mean of all predicted positions (including off-frame
extrapolations), and its weight equals the mean per-point visibility.
When CoTracker3 is unavailable, Pyramidal Lucas--Kanade optical flow
(21$\times$21 window, 3 pyramid levels, 30 iterations, $\varepsilon = 0.01$)
is used as a fallback; a tracked point is considered visible if its LK
status is~1, it lies within the image boundary, and its tracking residual
is below 30\,px.

\textbf{Signal C --- SAM2 Watchdog.}
SAM2 (\texttt{sam2.1\_hiera\_large}) is initialised with the lock logit and
the padded lock bounding box. On each subsequent frame, it predicts a mask
and centroid; its weight is
$w_C = \max\bigl(0,\,\min(0.4,\,s_{\text{SAM2}}\,(1 - r_{\text{border}}))
\bigr)$,
where $s_{\text{SAM2}}$ is the prediction score and $r_{\text{border}}$
penalises masks that touch the image border. Signal~C acts primarily as a
watchdog: if it disagrees with the fused estimate by more than 60\,px for
five consecutive frames, an alarm is raised.

\textbf{Signal D --- DINOv2 Best-Buddy Matching.}
DINOv2 (\texttt{dinov2\_vits14}, ViT-S/14, patch size 14) embeds both the
lock crop and the current frame into patch-level feature maps
(L2-normalised). Bidirectional best-buddy matching retains a patch pair
$(i,j)$ only if $j = \arg\max_k \langle f_i^{\text{ref}}, f_k^{\text{cur}}\rangle$
and $i = \arg\max_k \langle f_j^{\text{cur}}, f_k^{\text{ref}}\rangle$
simultaneously. Pairs with cosine similarity below 0.55 or when fewer than 8
bidirectional matches survive are discarded. The signal centroid is the
similarity-weighted mean of the matched patch centres in the current frame,
and $w_D = \bar{s}_D \cdot \min(1,\,n/30)$, where $\bar{s}_D$ is the mean
matched similarity and $n$ the number of valid matches.

\textbf{Fusion.}
The four signal estimates and weights are combined via the Weiszfeld weighted
geometric median (12 Weiszfeld iterations, convergence threshold
$\varepsilon = 10^{-3}$). If Signal~A is available and the fused result
deviates from Signal~A by more than 80\,px, the fused estimate is overridden
by Signal~A.

\subsubsection{TERMINAL State}
\label{sec:terminal}

A transition to the \textsc{terminal} state is triggered when either the
depth at the fused centroid drops below $30$\,mm or the fused pixel error
$\lVert\mathbf{e}\rVert_2 < 8$\,px. In the terminal state, a RANSAC plane
fit is applied to a $100\times100$\,px depth patch centred on the fused
centroid (100 iterations, 4\,mm inlier threshold). The fitted plane normal
provides the final approach direction for contact.


% =========================================================================
\section{Baseline Methods for Ablation}
\label{sec:baselines}

To evaluate the contribution of each component, the following baseline
variants are compared against the proposed SAM3 pipeline.

\subsection{GroundingDINO + SAM2}
\label{sec:gdino}

This baseline, implemented in \texttt{servo\_pipeline.py} with
\texttt{DETECTOR=gdino}, uses GroundingDINO for open-vocabulary bounding-box
detection followed by SAM2 for instance segmentation. The text prompt
``\texttt{box}'' is expanded to ``\texttt{box. object. thing.}'' to
improve recall. Detection thresholds are $\tau_{\text{box}} = 0.25$ and
$\tau_{\text{text}} = 0.25$; a permissive retry at $0.15$ is executed when
the first pass returns no detections.

SAM2 (\texttt{sam2.1\_hiera\_large}) is conditioned on the selected bounding
box and its geometric centre as a positive point prompt, with
\texttt{multimask\_output=False}. Full GroundingDINO redetection runs every
30 frames; between redetections, the prior bounding box is forwarded as a
SAM2 input without re-querying GroundingDINO.

\subsection{OWLv2 + SAM2}
\label{sec:owlv2}

A second variant in \texttt{servo\_pipeline.py} replaces GroundingDINO with
OWLv2 (\texttt{DETECTOR=owlv2}), which supports an optional image-guided
detection mode: the reference crop is passed alongside the text prompt to
bias detection toward instances visually similar to the reference.
The SAM2 segmentation stage and servo control law are identical to the
GroundingDINO baseline.

\subsection{DINOv2 with Color Similarity}
\label{sec:dinov2color}

Implemented in \texttt{variants/dinov2\_servo\_color.py}, this variant
replaces the text-prompted detector with DINOv2 patch-feature matching
augmented by colour similarity. Candidate patches are ranked by a combined
score of DINOv2 cosine similarity and colour histogram overlap
($w_{\text{color}} = 0.7$ for colour). The 93rd-percentile similarity
threshold is used for acceptance. SAM2 segmentation and the same servo
control law are retained.

\subsection{DINOv2 with SAM2 Logit Propagation}
\label{sec:dinov2sam2track}

Implemented in \texttt{variants/sam2\_tracking\_opencv.py}, this variant
separates a \textsc{detect} phase (DINOv2 ViT-B/14 + SAM2 single-frame
prediction) from a \textsc{track} phase (SAM2 logit propagation across
consecutive frames without re-running DINOv2). Redetection is triggered
when the SAM2 score drops below a threshold, the mask area changes by more
than a factor of three, or a fixed number of frames have elapsed. The servo
control law is unchanged.


% =========================================================================
\section{Experiments}
\label{sec:experiments}

\subsection{Hardware Setup}
\label{sec:hardware}

All experiments are conducted on a UFacory xArm (6-DOF) mounted to a
fixed tabletop, connected at IP \texttt{192.168.1.241}.  A ZED Mini stereo
camera is mounted rigidly on the tool flange in an eye-in-hand configuration.
Images are captured at HD720 resolution ($1280\times720$ pixels, 30\,fps).
When the ZED SDK (PyZED) is available, camera frames are obtained via the
native PyZED interface and raw SVO2 recordings are saved automatically.
When PyZED is unavailable, the left half of the $2560\times720$
OpenCV stereo frame is used and lens distortion is corrected using the ZED
factory calibration parameters loaded from the on-disk serial-number
configuration file.

\subsection{Ground Truth Measurement Protocol}
\label{sec:gt}

Alignment error is measured using an ArUco marker~\cite{} placed at the
geometric centre of the target box's top face. The marker dictionary is
\texttt{DICT\_4X4\_50} with marker side length 30\,mm. The ground-truth
procedure is as follows.

\subsubsection{Phase 1: Ground Truth Capture}

The \emph{marked} box (with the ArUco marker affixed) is placed at the
target location. The xArm is moved to the trial starting pose. A burst of
five consecutive ZED frames is captured; ArUco corners are detected in each
frame using \texttt{cv2.aruco.ArucoDetector} and the 3-D marker pose is
estimated via \texttt{cv2.aruco.estimatePoseSingleMarkers} with the ZED
factory intrinsics $(f_x, f_y, c_x, c_y)$. The marker translation vectors
are averaged across detected frames to reduce single-frame noise, then
transformed to the robot base frame:
\begin{equation}
  \mathbf{p}^{\text{gt}}_{\text{base}}
  = \mathbf{T}_{\text{base,ee}}\,\mathbf{T}_{\text{ee,cam}}\,
    \tilde{\mathbf{p}}^{\text{gt}}_{\text{cam}},
  \label{eq:gt_transform}
\end{equation}
where $\mathbf{T}_{\text{base,ee}}$ is the live forward-kinematics pose at
the moment of capture.

\subsubsection{Phase 2: Servoing Trial}

The marked box is replaced by a physically identical \emph{unmarked} box
placed in the same footprint. The xArm is returned to the trial starting pose.
The SAM3 servo pipeline (\texttt{servo\_pipeline\_sam3\_exp1.py}) is launched
with \texttt{--convergence-output} set to the trial output directory.
The operator enables servoing by pressing \textbf{v} in the OpenCV display
window. The pipeline runs until convergence is detected: the radial pixel
error $\lVert\mathbf{e}\rVert_2 \leq 15$\,px (the dead zone) for five
consecutive servo steps. Upon convergence, the final EE pose is read from
the xArm SDK (\texttt{arm.get\_position()}) and written atomically to
\texttt{final\_position.json}; the pipeline then exits automatically.

\subsection{Evaluation Metrics}
\label{sec:metrics}

From the pair $(\mathbf{p}^{\text{gt}}_{\text{base}},
\mathbf{p}^{\text{final}}_{\text{base}})$, three alignment error components
are computed:
\begin{align}
  \varepsilon_{3D} &= \lVert \mathbf{p}^{\text{final}} -
                             \mathbf{p}^{\text{gt}} \rVert_2, \\
  \varepsilon_{\text{lat}} &= \sqrt{(\Delta y)^2 + (\Delta z)^2}, \\
  \varepsilon_{\text{depth}} &= |\Delta x|,
\end{align}
where $x$ is the forward axis (toward the box), $y$ and $z$ are the lateral
axes in the xArm base frame. Lateral error $\varepsilon_{\text{lat}}$
quantifies planar alignment of the EE above the box centre.

Convergence time $t_{\text{conv}}$ is the wall-clock duration from the first
enabled servo step to the fifth consecutive in-zone frame. Path length
$\ell$ is the cumulative Euclidean distance of the EE trajectory
$\sum_{i=1}^{N-1} \lVert \mathbf{p}_i - \mathbf{p}_{i-1} \rVert_2$.

\subsection{Trial Design}
\label{sec:trial_design}

The experiment follows a full factorial design:
\begin{center}
\begin{tabular}{ll}
  Box types ($\mathcal{B}$):    & B1--B5 \quad (5 types)  \\
  Configurations ($\mathcal{C}$): & C1--C4 \quad (4 configs) \\
  Starting poses ($\mathcal{P}$): & P1--P3 \quad (3 poses)  \\
  Trials per cell:              & 10                        \\
  \midrule
  Total trials:                 & $5\times4\times3\times10 = 600$
\end{tabular}
\end{center}

Box types are summarised in Table~\ref{tab:boxes};
configurations in Table~\ref{tab:configs};
starting poses in Table~\ref{tab:poses}.

\begin{table}[h]
  \centering
  \caption{Box Types}
  \label{tab:boxes}
  \begin{tabular}{@{}lll@{}}
    \toprule
    ID & Description & Test purpose \\
    \midrule
    B1 & Plain brown cardboard             & Baseline, minimal texture \\
    B2 & Printed graphics on all faces     & Confusion with surface print \\
    B3 & Glossy / laminated surface        & Specular reflection \\
    B4 & Dark (black or dark blue)         & Low-contrast segmentation \\
    B5 & Small form factor (phone-sized)   & Small-target precision \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}[h]
  \centering
  \caption{Stack Configurations}
  \label{tab:configs}
  \begin{tabular}{@{}lll@{}}
    \toprule
    ID & Description         & Stack height \\
    \midrule
    C1 & Single box on table & 1 \\
    C2 & Two identical boxes & 2 \\
    C3 & Three identical boxes & 3 \\
    C4 & Two different boxes (target on top) & 2 \\
    \bottomrule
  \end{tabular}
\end{table}

\begin{table}[h]
  \centering
  \caption{Starting Poses}
  \label{tab:poses}
  \begin{tabular}{@{}lll@{}}
    \toprule
    ID & Description                            & Approx.\ standoff \\
    \midrule
    P1 & Centred above, large standoff          & $\sim$400\,mm \\
    P2 & Left-offset, medium standoff           & $\sim$200\,mm \\
    P3 & Right-offset, close standoff           & $\sim$150\,mm \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Results}
\label{sec:results}

Table~\ref{tab:main_results} reports the primary alignment and timing metrics
aggregated over 10 trials per cell (starting poses pooled). Values will be
filled after data collection.

\begin{table*}[t]
  \centering
  \caption{Table~I: Servoing Precision Across Box Types and Configurations
           (mean $\pm$ std, $N=10$ per cell, starting poses pooled).
           $\varepsilon_{\text{lat}}$: lateral error (mm);
           $\varepsilon_{\text{depth}}$: depth error (mm);
           $\varepsilon_{3D}$: 3-D Euclidean error (mm);
           $t_{\text{conv}}$: convergence time (s).}
  \label{tab:main_results}
  \begin{tabular}{@{}llcccc@{}}
    \toprule
    Box & Config
      & $\varepsilon_{\text{lat}}$\,(mm)
      & $\varepsilon_{\text{depth}}$\,(mm)
      & $\varepsilon_{3D}$\,(mm)
      & $t_{\text{conv}}$\,(s) \\
    \midrule
    B1 & C1 & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    B1 & C2 & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    B1 & C3 & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    B1 & C4 & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    B2 & C1 & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    B2 & C2 & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    B2 & C3 & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    B2 & C4 & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    B3 & C1 & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    B3 & C2 & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    B3 & C3 & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    B3 & C4 & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    B4 & C1 & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    B4 & C2 & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    B4 & C3 & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    B4 & C4 & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    B5 & C1 & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    B5 & C2 & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    B5 & C3 & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    B5 & C4 & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    \midrule
    ALL & ALL & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- & ---\,$\pm$\,--- \\
    \bottomrule
  \end{tabular}
\end{table*}

Table~\ref{tab:perception_results} reports per-condition perception quality
metrics averaged over all trials.

\begin{table*}[t]
  \centering
  \caption{Table~II: Perception Metrics Per Condition
           (mean over all trials per cell).
           Det.\ Recall: fraction of servo frames with at least one valid
           SAM3 detection;
           Sim.: mean ResNet-18 cosine similarity between chosen crop and
           reference image.}
  \label{tab:perception_results}
  \begin{tabular}{@{}llcc@{}}
    \toprule
    Box & Config & Det.\ Recall\,(\%) & Similarity \\
    \midrule
    B1 & C1 & --- & --- \\
    B1 & C2 & --- & --- \\
    B2 & C1 & --- & --- \\
    B2 & C2 & --- & --- \\
    B3 & C1 & --- & --- \\
    B3 & C2 & --- & --- \\
    B4 & C1 & --- & --- \\
    B4 & C2 & --- & --- \\
    B5 & C1 & --- & --- \\
    B5 & C2 & --- & --- \\
    \midrule
    ALL & ALL & --- & --- \\
    \bottomrule
  \end{tabular}
\end{table*}


% =========================================================================
\section{Open Questions and Assumptions}
\label{sec:assumptions}

The following assumptions and unresolved technical points are identified in
the current implementation.

\subsection{Hand-Eye Calibration}

Equations~\eqref{eq:transform} and~\eqref{eq:gt_transform} require the
hand-eye matrix $\mathbf{T}_{\text{ee,cam}}$ relating the ZED Mini optical
frame to the xArm tool-centre-point frame. This calibration has not yet been
performed. In the current implementation, $\mathbf{T}_{\text{ee,cam}}$ defaults
to the $4\times4$ identity matrix, which is valid only if the camera frame
and tool frame are co-located and co-aligned. Any physical offset or rotation
of the camera mount introduces a systematic bias in ground-truth
measurements~\eqref{eq:gt_transform} and disables Signal~A in the last-mile
fusion (Section~\ref{sec:lastmile}). Hand-eye calibration must be performed
before quantitative 3-D alignment results can be reported.

\subsection{Depth Source for the Last-Mile Phase}

The last-mile pipeline (\texttt{servo\_lastmile.py}) relies on metric
depth values in millimetres to evaluate the lock-transition depth criterion
($150 \leq z \leq 400$\,mm) and the terminal condition ($z < 30$\,mm), as
well as to back-project the centroid to 3-D via~\eqref{eq:backproject}.
When PyZED is available, metric stereo depth is provided directly by the
ZED SDK. The far-range SAM3 pipeline also uses Depth-Anything-V2
(ViT-Small encoder) for depth inference; however, Depth-Anything-V2
produces relative, non-metric depth values that cannot be used directly
for the millimetre-scale thresholds in the last-mile state machine
without additional scale calibration.

\subsection{xArm Base Frame Axis Convention}

The alignment error decomposition in Section~\ref{sec:metrics} assigns the
$x$-axis to the forward direction (toward the box), and $y$, $z$ to the
lateral plane. The xArm SDK reports EE positions in its own base frame whose
axis convention must be verified empirically for the specific table mounting
used in experiments.

\subsection{Two-Box Swap Ground Truth Error Floor}

The ground truth protocol (Section~\ref{sec:gt}) replaces the marked box
with a physically identical unmarked box. Any positional error in placing the
unmarked box in the same footprint as the marked box contributes a systematic
offset to the measured alignment error. This placement repeatability error
constitutes the ground-truth error floor and should be characterised
independently before reporting results.

\subsection{Integration of Far-Range and Last-Mile Phases}

In the current codebase, the SAM3 far-range pipeline
(\texttt{servo\_pipeline\_sam3.py}) and the last-mile multi-signal fusion
pipeline (\texttt{servo\_lastmile.py}) are implemented as separate scripts.
The experiment in Section~\ref{sec:experiments} evaluates only the SAM3
far-range phase. Integration of the two phases into a single continuous
pipeline (with an automatic handoff at the LOCK transition) is a planned
but not yet implemented step.

\subsection{Servo Update Rate and SAM3 Latency}

The minimum inter-command interval is $0.3$\,s, giving a maximum control
rate of approximately $3.3$\,Hz. SAM3 is queried on every camera frame
(REDETECT\_INTERVAL\,=\,1), meaning that the effective control rate is
limited by SAM3 inference time in addition to the $0.3$\,s floor. On
the GPU used in experiments (unspecified at time of writing), SAM3 inference
latency has not yet been characterised.


% =========================================================================
\section*{Acknowledgments}
I would like to thank Prof.\ Jeffrey Ichnowski for his invaluable guidance,
technical insight, and consistent support throughout this project. His
mentorship at the Momentum Robotics Lab has been central in shaping both the
research direction and the system design presented in this work.
The author also thanks the members of the Momentum Robotics Lab for their
helpful discussions and feedback during development. Access to the xArm
manipulator, ZED Mini stereo camera, and associated lab infrastructure was
essential to the real-world experiments reported here.
This work was carried out as part of an independent study in the Master of
Science in Computer Vision program at Carnegie Mellon University's Robotics
Institute.

\vfill

\end{document}
